### 2016-06-01 pics/ex_complete.tex

\documentclass{standalone}
\usepackage{tikz}
\begin{document}
	\begin{tikzpicture}

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (0,0) (1) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{1,3}$] at (1,-0.5) (2) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (2,0) (3) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,4}$] at (1,-1) (4) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,1}$] at (1,1) (5) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{1,2}$] at (1,0.5) (6) {};

		\path (1)
		edge (2)
		edge (4)
		edge (5)
		edge (6)
		(3)
		edge (2)
		edge (4)
		edge (5)
		edge (6);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (0,-3) (7) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{2,3}$] at (1,-3.5) (8) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (2,-3) (9) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,4}$] at (1,-4) (10) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,1}$] at (1,-2) (11) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{2,2}$] at (,-2.5) (12) {};

		\path (7)
		edge (8)
		edge (12)
		edge (10)
		edge (11)
		(9)
		edge (8)
		edge (10)
		edge (11)
		edge (12);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (0,-6) (12) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{3,3}$] at (1,-6.5) (14) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (2,-6) (15) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,4}$] at (1,-7) (16) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,1}$] at (1,-5) (17) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{3,2}$] at (1,-5.5) (18) {};

		\path (12)
		edge (14)
		edge (16)
		edge (17)
		edge (18)
		(15)
		edge (14)
		edge (16)
		edge (17)
		edge (18);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (0,-9) (19) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{4,3}$] at (1,-9.5) (20) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (2,-9) (21) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,4}$] at (1,-10) (22) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,1}$] at (1,-8) (23) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{4,2}$] at (1,-8.5) (24) {};

		\path (19)
		edge (20)
		edge (22)
		edge (23)
		edge (24)
		(21)
		edge (20)
		edge (22)
		edge (23)
		edge (24);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,2}$] at (2,2) (d1) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,3}$] at (3,1) (d2) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,4}$] at (4,1) (d3) {};

		\path[orange] (d1)
		edge (7)
		edge (3)
		(d2)
		edge (12)
		edge (3)
		(d3)
		edge (19)
		edge (3);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,1}$] at (0,-1) (d4) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,4}$] at (3,-2) (d5) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,3}$] at (4,-3) (d6) {};

		\path[green] (d4)
		edge (1)
		edge (9)
		(d5)
		edge (19)
		edge (9)
		(d6)
		edge (12)
		edge (9);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,1}$] at (3,-4) (d7) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,2}$] at (-1,-4) (d8) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,4}$] at (-1,-5) (d9) {};

		\path[red](d7)
		edge (1)
		edge (15)
		(d8)
		edge (7)
		edge (15)
		(d9)
		edge (19)
		edge (15);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,1}$] at (-1,-7) (d10) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,2}$] at (4,-8) (d11) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,3}$] at (-1,-8) (d12) {};

		\path[blue] (d10)
		edge (1)
		edge (21)
		(d11)
		edge (7)
		edge (21)
		(d12)
		edge (12)
		edge (21);

		\node[draw,circle,inner sep=0.05cm, fill, label=right:c] at (5, -5) (c) {};
		\path (c)
		edge (21)
		edge (15)
		edge(3)
		edge(9);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (6, 0) (a11) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,1}$] at (6.5, 0) (c11) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (7, 0) (b11) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (6, -3) (a12) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,1}$] at (6.5, -3) (c12) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (7, -3) (b12) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (6, -6) (a13) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,1}$] at (6.5, -6) (c13) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (7, -6) (b13) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (6, -9) (a14) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,1}$] at (6.5, -9) (c14) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (7, -9) (b14) {};

		\path (c)
		edge (a11)
		edge (a12)
		edge (a13)
		edge (a14)
		(c11)
		edge (a11)
		edge (b11)
		(c12)
		edge (a12)
		edge (b12)
		(c13)
		edge (a13)
		edge (b13)
		(c14)
		edge (a14)
		edge (b14);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (8, 0) (a21) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,2}$] at (8.5, 0) (c21) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (9, 0) (b21) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (8, -3) (a22) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,2}$] at (8.5, -3) (c22) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (9, -3) (b22) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (8, -6) (a23) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,2}$] at (8.5, -6) (c23) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (9, -6) (b23) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (8, -9) (a24) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,2}$] at (8.5, -9) (c24) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (9, -9) (b24) {};

		\path (c21)
		edge (a21)
		edge (b21)
		(c22)
		edge (a22)
		edge (b22)
		(c23)
		edge (a23)
		edge (b23)
		(c24)
		edge (a24)
		edge (b24);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (10, 0) (a31) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,3}$] at (10.5, 0) (c31) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (11, 0) (b31) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (10, -3) (a32) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,3}$] at (10.5, -3) (c32) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (11, -3) (b32) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (10, -6) (a33) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,3}$] at (10.5, -6) (c33) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (11, -6) (b33) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (10, -9) (a34) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,3}$] at (10.5, -9) (c34) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (11, -9) (b34) {};

		\path (c31)
		edge (a31)
		edge (b31)
		(c32)
		edge (a32)
		edge (b32)
		(c33)
		edge (a33)
		edge (b33)
		(c34)
		edge (a34)
		edge (b34);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (12, 0) (a41) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,4}$] at (12.5, 0) (c41) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (13, 0) (b41) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (12, -3) (a42) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,4}$] at (12.5, -3) (c42) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (13, -3) (b42) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (12, -6) (a43) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,4}$] at (12.5, -6) (c43) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (13, -6) (b43) {};

		\node[draw,circle,inner sep=0.05cm, fill, label=above:a] at (12, -9) (a44) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,4}$] at (12.5, -9) (c44) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:b] at (13, -9) (b44) {};

		\path (c41)
		edge (a41)
		edge (b41)
		(c42)
		edge (a42)
		edge (b42)
		(c43)
		edge (a43)
		edge (b43)
		(c44)
		edge (a44)
		edge (b44);

		\node[draw,circle,inner sep=0.05cm, fill, label=right:c] at (14, -5) (c2) {};

		\path (c2)
		edge (b41)
		edge (b42)
		edge (b43)
		edge (b44);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{1,2}$] at (7,-1) (f1) {};

		\path[orange] (f1)
		edge (b11)
		edge (a22);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{2,1}$] at (7.25,-2) (f2) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$d_{2,3}$] at (7.5,-4) (f3) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,4}$] at (7,-4) (f4) {};

		\path[green] (f2)
		edge (b12)
		edge (a21)
		(f3)
		edge (b12)
		edge (a23)
		(f4)
		edge (b12)
		edge (a24);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{3,2}$] at (6.5,-4.5) (f5) {};

		\path[red] (f5)
		edge (b13)
		edge (a22);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{4,2}$] at (7,-7.5) (f6) {};

		\path[blue] (f6)
		edge (b14)
		edge (a22);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{1,2}$] at (9,-1) (f7) {};

		\path[orange] (f7)
		edge (b21)
		edge (a32);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{2,1}$] at (9.25,-2) (f8) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$d_{2,3}$] at (9.5,-4) (f9) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,4}$] at (9,-4) (f10) {};

		\path[green] (f8)
		edge (b22)
		edge (a31)
		(f9)
		edge (b22)
		edge (a33)
		(f10)
		edge (b22)
		edge (a34);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{3,2}$] at (8.5,-4.5) (f11) {};

		\path[red] (f11)
		edge (b23)
		edge (a32);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{4,2}$] at (9,-7.5) (f12) {};

		\path[blue] (f12)
		edge (b24)
		edge (a32);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{1,4}$] at (11,-1) (f13) {};

		\path[orange] (f13)
		edge (b31)
		edge (a42);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{2,1}$] at (11.25,-2) (f14) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$d_{2,3}$] at (11.5,-4) (f15) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,4}$] at (11,-4) (f16) {};

		\path[green] (f14)
		edge (b32)
		edge (a41)
		(f15)
		edge (b32)
		edge (a43)
		(f16)
		edge (b32)
		edge (a44);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{3,2}$] at (10.5,-4.5) (f17) {};

		\path[red] (f17)
		edge (b33)
		edge (a42);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:$d_{4,2}$] at (11,-7.5) (f18) {};

		\path[blue] (f18)
		edge (b34)
		edge (a42);
	\end{tikzpicture}
\end{document}