### 2016-06-01 pics/ex_con.tex

\documentclass{standalone}
\usepackage{tikz}
\begin{document}
	\begin{tikzpicture}

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (0,0) (1) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{1,3}$] at (0.5,1) (2) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (0,2) (3) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,4}$] at (1,1) (4) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{1,1}$] at (-1,1) (5) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{1,2}$] at (-0.5,1) (6) {};

		\path (1)
		edge (2)
		edge (4)
		edge (5)
		edge (6)
		(3)
		edge (2)
		edge (4)
		edge (5)
		edge (6);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (3,0) (7) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{2,3}$] at (3.5,1) (8) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (3,2) (9) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,4}$] at (4,1) (10) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{2,1}$] at (2,1) (11) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{2,2}$] at (2.5,1) (12) {};

		\path (7)
		edge (8)
		edge (12)
		edge (10)
		edge (11)
		(9)
		edge (8)
		edge (10)
		edge (11)
		edge (12);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (6,0) (12) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{3,3}$] at (6.5,1) (14) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (6,2) (15) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,4}$] at (7,1) (16) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{3,1}$] at (5,1) (17) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{3,2}$] at (5.5,1) (18) {};

		\path (12)
		edge (14)
		edge (16)
		edge (17)
		edge (18)
		(15)
		edge (14)
		edge (16)
		edge (17)
		edge (18);

		\node[draw,circle,inner sep=0.05cm, fill, label=left:a] at (9,0) (19) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$c_{4,3}$] at (9.5,1) (20) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=left:b] at (9,2) (21) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,4}$] at (10,1) (22) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$c_{4,1}$] at (8,1) (23) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=right:$c_{4,2}$] at (8.5,1) (24) {};

		\path (19)
		edge (20)
		edge (22)
		edge (23)
		edge (24)
		(21)
		edge (20)
		edge (22)
		edge (23)
		edge (24);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,2}$] at (1,3) (d1) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,3}$] at (2,3) (d2) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{1,4}$] at (1,2) (d3) {};

		\path[orange] (d1)
		edge (7)
		edge (3)
		(d2)
		edge (12)
		edge (3)
		(d3)
		edge (19)
		edge (3);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,1}$] at (1,-1) (d4) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,4}$] at (2,-1) (d5) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{2,3}$] at (2,0) (d6) {};

		\path[green] (d4)
		edge (1)
		edge (9)
		(d5)
		edge (19)
		edge (9)
		(d6)
		edge (12)
		edge (9);

		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,1}$] at (6,3) (d7) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,2}$] at (7,2) (d8) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=above:$d_{3,4}$] at (7,3) (d9) {};

		\path[red](d7)
		edge (1)
		edge (15)
		(d8)
		edge (7)
		edge (15)
		(d9)
		edge (19)
		edge (15);

		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,1}$] at (7,-1) (d10) {};
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,2}$] at (8,-1) (d11) {};		
		\node[draw,circle,inner sep=0.05cm, fill, label=below:$d_{4,3}$] at (8, 0) (d12) {};

		\path[blue] (d10)
		edge (1)
		edge (21)
		(d11)
		edge (7)
		edge (21)
		(d12)
		edge (12)
		edge (21);
	\end{tikzpicture}
\end{document}