### 2016-06-01 pics/k.tex

\documentclass{standalone}
\usepackage{tikz}
\begin{document}
	\begin{tikzpicture}

		\node[draw,circle,inner sep=0.05cm, fill] at (0,0) (1) {};
		\node[draw,circle,inner sep=0.05cm, fill] at (0,1) (2) {};
		\node[draw,circle,inner sep=0.05cm, fill] at (0,2) (3) {};
		\node[draw,circle,inner sep=0.05cm, fill] at (1,1) (4) {};
		\node[draw,circle,inner sep=0.05cm, fill] at (-1,1) (5) {};

		\path (1)
		edge (2)
		edge (4)
		edge (5)
		(3)
		edge (2)
		edge (4)
		edge (5);
	\end{tikzpicture}
\end{document}